### 2017-08-21 reports/thesis/book.tex
Added series evaluation to thesis.

\subsection{Saddlepoint approximation}
The m.g.f.~of a random variable $X$ is defined to be
\begin{equation}
M_X(\theta)=\expectation[\euler^{\theta X}]
=
\int_{-\infty}^{\infty}\euler^{x\theta} p_X(x) \diff x \ .
\end{equation}
For a given m.g.f., the saddlepoint approximation \cite{daniels1954saddlepoint} \cite{butler2007saddlepoint} finds an approximate $p_X(x)$. The saddlepoint approximation is given as
\begin{equation}
p_X(x)\approx\left(2\pi K_X''(s)\right)^{-1/2}\exp\left[K_X(s)-sx\right]
\label{eq:saddlePoint:generalSaddlePoint}
\end{equation}
where $K_X(\theta) = \ln\left(M_X(\theta)\right)$, and $s=s(x)$ is the solution to the saddle point equation $K_X'(s)=x$.

For the compound Poisson-Gamma distribution, the saddle point approximation is given as 
\begin{multline}
p_X(x)\approx
\frac{\left(\lambda\alpha\beta^\alpha\right)^{\frac{1}{2(\alpha+1)}}\euler^{-\lambda}}{\sqrt{2\pi(\alpha+1)}}x^{-\frac{\alpha+2}{2(\alpha+1)}}
\euler^{-x\beta}
\exp\left[x^{\frac{\alpha}{\alpha+1}}
	\frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
\right]
\\
\text{for }x>0 \ .
\label{eq:saddle_point_approx}
\end{multline}
The integral of the density approximation over the support may not equal to one, it can be numerically re-normalised if necessary. Thus it may be more sensible to write the approximation up to a constant
\begin{equation}
p_X(x)\propto x^{-\frac{\alpha+2}{2(\alpha+1)}}\euler^{-x\beta}\exp\left[x^{\frac{\alpha}{\alpha+1}}
	\frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
\right] \ .
\end{equation}

@@

>>>>

\subsection{Saddlepoint approximation}
The m.g.f.~of a random variable $X$ is defined to be
\begin{equation}
M_X(\theta)=\expectation[\euler^{\theta X}]
=
\int_{-\infty}^{\infty}\euler^{x\theta} p_X(x) \diff x \ .
\end{equation}
For a given m.g.f., the saddlepoint approximation \citep{daniels1954saddlepoint} \citep{butler2007saddlepoint} finds an approximate $p_X(x)$. The saddlepoint approximation is given as
\begin{equation}
p_X(x)\approx\left(2\pi K_X''(s)\right)^{-1/2}\exp\left[K_X(s)-sx\right]
\label{eq:saddlePoint:generalSaddlePoint}
\end{equation}
where $K_X(\theta) = \ln\left(M_X(\theta)\right)$, and $s=s(x)$ is the solution to the saddle point equation $K_X'(s)=x$.

For the compound Poisson-Gamma distribution, the saddle point approximation is given as 
\begin{multline}
p_X(x)\approx
\frac{\left(\lambda\alpha\beta^\alpha\right)^{\frac{1}{2(\alpha+1)}}\euler^{-\lambda}}{\sqrt{2\pi(\alpha+1)}}x^{-\frac{\alpha+2}{2(\alpha+1)}}
\euler^{-x\beta}
\exp\left[x^{\frac{\alpha}{\alpha+1}}
	\frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
\right]
\\
\text{for }x>0 \ .
\label{eq:saddle_point_approx}
\end{multline}
It appears the approximation is not well defined for $x=0$.

The integral of the density approximation over the support may not equal to one, it can be numerically re-normalised if necessary. Thus it may be more sensible to write the approximation up to a constant
\begin{equation}
p_X(x)\propto x^{-\frac{\alpha+2}{2(\alpha+1)}}\euler^{-x\beta}\exp\left[x^{\frac{\alpha}{\alpha+1}}
	\frac{(\lambda\beta^\alpha)^{\frac{1}{\alpha+1}}(\alpha+1)}{\alpha^{\frac{\alpha}{\alpha+1}}}
\right] \ .
\end{equation}

@@

\subsection{Series Evaluation}
Evaluation of the p.d.f.~of the compound Poisson-Gamma distribution is difficult because of the infinite sum
\begin{equation}
p_X(x) = 
\delta(x) \euler^{-\lambda}
+
\euler^{-\beta x-\lambda}\frac{1}{x}\sum_{y=1}^{\infty}W_y
\quad\text{for }x\geqslant 0
\end{equation}
where
\begin{equation}
W_y = \frac{\beta^{y\alpha}\lambda^yx^{y\alpha}}{\Gamma(y\alpha)y!} \ .
\end{equation}
$W_y$ can be expressed in a different way using different parameters
\begin{equation}
W_y(x,p,\phi)=\frac{x^{y\alpha}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
\end{equation}
where
\begin{equation}
p=\frac{2+\alpha}{1+\alpha}
\end{equation}
and
\begin{equation}
\phi = \frac{\alpha+1}{\beta^{2-p}(\lambda\alpha)^{p-1}} \ .
\end{equation}

\cite{dunn2005series} approximated the sum by truncation
\begin{equation}
\sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y
\end{equation}
where $y_\text{u}<y_{\text{max}}<y_\text{l}$ and $y_{\text{max}}$ is the value of $y$ which maximises $W_y$. \cite{dunn2005series} used Stirling's approximation to find $y_{\text{max}}$. This was done by treating $W_y$ as a continuous function of $y$ and is differentiable with respect to $y$.

It is easier to differentiate $\ln(W_y)$ where
\begin{multline*}
\ln(W_y) = y\alpha\ln(x)-y(1+\alpha)\ln(\phi)-y\alpha\ln(p-1)\\-y\ln(2-p)-\ln(y!)-\ln\Gamma(y\alpha)
\end{multline*}
\begin{equation}
\ln(W_y) =
y\ln\left(
	\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)}
\right)
-\ln(y!)-\ln\Gamma(y\alpha) \ .
\end{equation}
Using Stirling's approximation
\begin{equation}
\ln(n!)\approx\ln\Gamma(n)\approx n\ln(n!)-n
\end{equation}
then
\begin{equation}
\ln(W_y) \approx
y\ln\left(
	\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)}
\right)
-y\ln(y)+y-y\alpha\ln(y\alpha) + y\alpha \ .
\end{equation}
Taking the derivative with respect to $y$
\begin{multline*}
\frac{\partial \ln(W_y)}{\partial y} \approx
\ln\left(
	\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)}
\right)
-\ln(y)-1+1
\\-\alpha\ln(y\alpha)-\alpha+\alpha
\end{multline*}
\begin{equation}
\frac{\partial \ln(W_y)}{\partial y} \approx
\ln\left(
	\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)}
\right)
-\ln(y)
-\alpha\ln(y\alpha) \ .
\end{equation}

Setting the derivative to zero
\begin{equation*}
0 \approx \ln\left(
\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)y_{\text{max}}^{1+\alpha}\alpha^\alpha}
\right)
\end{equation*}
\begin{equation*}
1 \approx 
\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)y_{\text{max}}^{1+\alpha}\alpha^\alpha}
\end{equation*}
\begin{equation*}
y_{\text{max}}^{1+\alpha} \approx 
\frac{x^\alpha}{\phi^{1+\alpha}(p-1)^\alpha(2-p)\alpha^\alpha}
\end{equation*}
\begin{equation*}
y_{\text{max}} \approx 
\frac{1}{\phi}
\left(
	\frac{x}{(p-1)\alpha}
\right)^{\frac{\alpha}{1+\alpha}}
(2-p)^{\frac{-1}{1+\alpha}} \ .
\end{equation*}
This can be simplified using the fact that 
\begin{equation*}
\alpha=\frac{2-p}{p-1} \ ,
\end{equation*}
\begin{equation*}
\frac{1}{1+\alpha} = p-1
\end{equation*}
and
\begin{equation*}
\frac{\alpha}{1+\alpha} = 2-p
\end{equation*}
then
\begin{equation*}
y_{\text{max}} \approx 
\frac{1}{\phi}
\left(
	\frac{x}{2-p}
\right)^{2-p}
(2-p)^{1-p}
\end{equation*}
and finally
\begin{equation}
y_{\text{max}} \approx \frac{x^{2-p}}{\phi(2-p)} \ .
\end{equation}
Because values of $y$ should be a positive integer, it would be appropriate to round the solution to $y_\text{max}$ accordingly
\begin{equation}
y_{\text{max}} = \text{max}\left[
1,\text{round}\left(\frac{x^{2-p}}{\phi(2-p)}\right)
\right] \ .
\end{equation}

To verify that $y_\text{max}$ is a maximum, the second derivative can be investigated
\begin{equation}
\frac{\partial^2\ln(W_y)}{\partial y^2}
\approx
-\frac{1}{y}(\alpha+1)
\end{equation}
to see that
\begin{equation}
\frac{\partial^2\ln(W_y)}{\partial y^2} < 0 \quad \text{for }y=1,2,3,\dotdotdot
\end{equation}
therefore $y_\text{max}$ is a maximum.

Returning back to the truncation of the infinite sum
\begin{equation}
\sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y \ ,
\end{equation}
$y_\text{l}$ and $y_\text{u}$ can be chosen such that $W_{y_\text{l}}$ and $W_{y_\text{u}}$ are less than $\epsilon W_{y_\text{max}}$ where $\epsilon$ is some small constant. To prevent overflow problems, it is advised to calculate each term in the summation in log scale and using the following equation
\begin{equation}
\ln\left[\sum_{y=y_\text{l}}^{y_\text{u}}W_y\right]
= \ln\left(W_{y_\text{max}}\right)+\ln\sum_{y=y_\text{l}}^{y_\text{u}}\exp\left[
\ln\left(W_y\right)-\ln\left(W_{y_\text{max}}\right)
\right] \ .
\end{equation}
